Space-time adaptivity for parabolic evolution equations

### 2020-05-29 Latex/paper.tex
new version

\usepackage{showlabels}
\usepackage[boxed]{algorithm2e}
\usepackage{amsaddr}
\usepackage{babel}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{commath}
\usepackage{colortbl}
\usepackage{dsfont}
\usepackage{enumerate}
\usepackage{esint}
\usepackage[left=0.09\paperwidth, right=0.15\paperwidth, bottom=0.12\paperheight]{geometry}
\usepackage{graphicx}
\usepackage[displaymath, mathlines]{lineno}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{multirow,bigdelim}
\usepackage{url}
\usepackage{pdflscape}
\usepackage{rotating}
%\usepackage[outer]{showlabels}
\usepackage[color=white,linecolor=black, textwidth=0.12\paperwidth]{todonotes}
%\setlength{\parindent}{0pt}
\usepackage{wrapfig}

@@

\usepackage{color}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\providecommand\C{}
\renewcommand{\C}{\mathbb{C}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\K}{\mathcal{K}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\bbD}{\mathbb{D}}
\newcommand{\f}{\varphi}
\newcommand{\e}{\varepsilon}
\newcommand{\eps}{\varepsilon}
\renewcommand{\a}{\alpha}
\renewcommand{\b}{\beta}
\newcommand{\z}{\zeta}
\renewcommand{\O}{\Omega}
\renewcommand{\P}{\mathbb P}
\newcommand{\intRn}{\int_{\R^n}}
\newcommand{\sumi}{\sum_{i=1}^n}
\newcommand{\sumij}{\sum_{i,j=1}^n}
\renewcommand{\d}{\delta}
\newcommand{\p}{\partial}
\DeclareMathOperator{\osc}{osc}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\vol}{vol}
\DeclareMathOperator{\gen}{gen}
\DeclareMathOperator{\cat}{Cat}
\DeclareMathOperator{\Cat}{Cat}
\DeclareMathOperator{\Div}{div}
\DeclareMathOperator{\blockdiag}{blockdiag}
\renewcommand{\div}{\Div}
\DeclareMathOperator{\spann}{span}
\DeclareMathOperator{\diam}{diam}
\DeclareMathOperator{\dist}{dist}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\usepackage{stmaryrd}
\newcommand{\jump}[1]{\left\llbracket #1 \right\rrbracket}
\newcommand{\rhu}{\rightharpoonup}
\newcommand{\T}{\mathcal{T}}
\newcommand{\V}{\mathcal{V}}
\DeclareMathOperator*{\dof}{\#\!DoF}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\Id}{Id}
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\definecolor{ao(english)}{rgb}{0.0, 0.5, 0.0}
\providecommand\note{}
\newcommand{\new}[1]{{\color{ao(english)} #1}}
\renewcommand{\L}{\mathcal L}
\newcommand{\Y}{\mathcal Y}
\newcommand{\Lto}{{L_2(\Omega)}}
\newcommand{\bbT}{\mathbb{T}}
\newcommand{\cL}{\mathcal L}
\newcommand{\Lis}{\cL\mathrm{is}}
\newcommand{\udelta}{{\underline{\delta}}}
\newcommand{\jw}[1]{{\color{red}{JW: #1}}}
\newcommand{\rvv}[1]{{\color{teal}{RvV: #1}}}
\DeclareMathOperator*{\ran}{ran}
\SetKwRepeat{Do}{do}{while}%

@@

\section{Introduction}
\section{Problem statement and solution}
Let $V$ and $H$ be separable Hilbert spaces on a spatial domain $\Omega$ such that
$V \hookrightarrow H$ with dense and compact embedding. Identifying $H$ with its
dual, we obtain the Gelfand triple $V \hookrightarrow H \simeq H' \hookrightarrow V'$.

@@

\begin{lemma}
  \label{lem:energy-norm}
  It holds that $\|\cdot\|_X^2=(S\cdot)(\cdot)$.
\end{lemma}
\begin{proof}
  We find \jw{dit ben ik nog niet nagegaan}
  \[
    \|w\|_X^2=\sup_{0 \neq v_1 \in Y} \frac{(B w)(v_1)}{\|v_1\|_Y^2}+\|\gamma_0 w\|^2
    =\sup_{0 \neq (v_1,v_2) \in Y \times H} \frac{((B w)(v_1)+\la \gamma_0 w,v_2\ra)^2}{\|v_1\|_Y^2+\|v_2\|^2} =(Sw)(w),
  \]
  where the first equality can be found in e.g.~\cite[Thm.~2.1]{Ern2017a}, and,
  when realising that $S$ is also the Schur complement of
  $\begin{bsmallmatrix} A_s & 0 & B\\ 0 & \Id & \gamma_0\\ B' & \gamma_0' & 0\end{bsmallmatrix} \in \cL(Y\times H \times X, Y'\times H'\times X')$, the last one follows from e.g.~\cite[Lemma~2.2]{Kondratyuk2008}.
\end{proof}

@@

Let us first argue why uniform stability for general nonuniform prismatic partitions
is out of reach. \jw{follows from $\dim Y^\delta \lesssim \dim X^\delta$ en
de non-locality van $A_s^{-1}$: gegeven een $X^\delta$ die naar een spacetime-hoekje
gerefined is, moet $Y^\delta$ de gehele time-slab refinen. Ik ben het precieze argument
even vergeten maar met een plaatje wordt het snel duidelijk.}

@@

\begin{prop} \label{prop:space-infsup}
  Let $W$ and $\tilde W$ be closed subspaces of $H$ such that there exists a
  (biorthogonal) projector $Q \in \cL(H,H)$ with $\ran Q=W$ and $\ran (\Id -Q)={\tilde W}^\perp$.
  Then, when $W \subset V$, it holds that
  \[
    \gamma:=\inf_{0 \neq \tilde w \in \tilde W}\sup_{0 \neq w \in W}\frac{\la \tilde w,w\ra}{\|\tilde w\|_{V'}\|w\|_V}>0
  \]
  if and only if $Q \in \cL(V,V)$, in which case $\gamma=\|Q\|_{\cL(V,V)}^{-1}$.
\end{prop}
\begin{proof} \jw{Dit ben ik nog niet nagegaan.} If $Q\in \cL(V,V)$, then for $\tilde w \in \tilde W$,
$$
\sup_{0 \neq w \in W}\frac{\la \tilde w,w\ra}{\|w\|_V}=\sup_{0 \neq v \in V}\frac{\la \tilde w, v\ra}{\|Qv\|_V}
\geq \|Q\|_{\cL(V,V)}^{-1} \|\tilde w\|_{V'},
$$
or $\gamma \geq \|Q\|_{\cL(V,V)}^{-1}$. If on the other hand $\gamma>0$, then for any $u \in H$,
$$
\gamma \|Q' u\|_{V'} \leq \sup_{0 \neq w \in W}\frac{\la Q'u,w\ra}{\|w\|_V} =
 \sup_{0 \neq w \in W}\frac{\la u,w\ra}{\|w\|_V} \leq \|u\|_{V'},
 $$
so that from $H$ being dense in $V'$, $\|Q\|_{\cL(V,V)}=\|Q'\|_{\cL(V',V')}\leq \gamma^{-1}$.
\end{proof}
\begin{remark}
  In the typical case where $H = L_2(\Omega)$ and $V = H_0^1(\Omega)$ on a bounded
  polytopal domain $\Omega \subset \R^d$, the uniform boundedness of $\|Q\|_{\cL(V,V)}$
  has been demonstrated for families of locally refined partitions, for $d=2$
  including those that are generated by the newest vertex bisection algorithm.
  We refer to \cite{Carstensen2001a,Gaspoz2016}. \jw{Misschien is het ``$H^1$-stability
  of $L^2$-projection" inmiddels wel uitgespeeld. ik vond ook een werk uit
  2014 van de groep van Dirk, en dit werkte voor elke dimensie, maar enkel op
  vreemde fin.elts. Ook een werk uit 2019 van Gaspoz, Heine, Siebert waar het
  2d-NVB geval wordt bekeken voor Lagrange-fin elts tot graad 4.}
\end{remark}

@@

In the spatial axis, we consider the collection $\mathbb T$ of conforming
triangulations generated by NVB from some coarsest triangulation $\T_\perp$,
and define $V_\T = \tilde V_\T$ the space of continuous piecewise linears wrt.~$\T$
that vanish on $\partial \Omega$ such that $X^\delta := \sum_{\lambda \in \vee^\delta_\Sigma} \sigma_\lambda \otimes V_{\T_{\vee^\delta_\Sigma, \lambda}}$ satisfies the doubletree constraint.
This means that when $\mu \in \vee^\delta_\Sigma$ is a parent of $\lambda \in \vee^\delta_\Sigma$,
    then $\T_{\vee^\delta_\Sigma, \mu}$ is a refinement of $\T_{\vee^\delta_\Sigma, \lambda}$, and that $\vee^\delta_\Sigma$ is a subtree of $\vee_\Sigma$, meaning
\begin{itemize}
  \item it contains all wavelets on level 0, and
  \item if it contains a wavelet of level $\lambda \geq 1$, then it also contains
    its parents.
\end{itemize}

@@

We consider spaces $Y^\delta = \tilde Y^\delta$ of the form
$Y^\delta := \sum_{\lambda \in \vee^\delta_\Psi} \psi_\lambda \otimes V_{\T_{\vee^\delta_\Psi, \lambda}}$
for $\T_{\vee^\delta_\Psi, \lambda} \in \mathbb T$, again satisfying the doubletree
constraint. The choice
\begin{equation}
  Y^\delta = \sum_{\lambda \in \vee_\Sigma^\delta} \sum_{\{\mu \in \vee_\Psi\colon |\mu|=|\lambda|,\,
  |\supp \sigma_\lambda \cap \supp \psi_\mu|>0\}} \psi_\mu \otimes V_{\T_{\vee_\Sigma^\delta,\lambda}}
  \label{eqn:generate-Ydelta}
\end{equation}
is a doubletree by virtue of $X^\delta$ being one, and for this pair $(X^\delta, Y^\delta)$,
\eqref{eqn:d_tX-in-Y} and \eqref{eqn:inclusion} are valid.

@@

\begin{assum}[$X$-stable splitting of $X^{\udelta} \ominus X^\delta$]
  \jw{is dit een juiste naam voor deze assumption?}
  Assume we have $\Psi^\delta=\{\psi^\delta_\lambda\colon \lambda \in I_\delta^\udelta\} \subset X^\udelta$
  such that $X^\delta + \spann \Psi^\delta=X^\udelta$ and, for some constants $0<m\leq M$,
  for all $z \in X^\delta$ and ${\bf c} := (c_\lambda)_{\lambda \in I_\delta^\udelta} \in \R^{\# I_\delta^\udelta}$, we have
  \begin{equation}
    m^2 \|z+{\bf c}^\top \Psi^\delta\|_X^2
    \leq \|z\|_X^2+\|{\bf c}\|^2
    \leq M^2 \|z+{\bf c}^\top \Psi^\delta\|_X^2.
  \label{eqn:X-stable-splitting}
  \end{equation}
  \label{assum:X-stable-splitting}
\end{assum}
\begin{remark}
  A suitable basis is not currently known to us. However, in \S TODO, we will construct
  such $\Psi^\delta$ with constants $m$ and $M$ that deteriorate with increasing $|\lambda|$.
  \jw{is die constructie iets dat we misschien in een appendix willen zetten?}
\end{remark}

@@

This error estimator will drive our adaptive loop. Even without Assumption~\ref{assum:X-stable-splitting} however, we derive the following efficient and reliable \emph{global estimator}.
\begin{prop}
  \label{prop:global-apost}
  Let $\Xi^{\udelta}$ be a basis for $Y^{\udelta}$. Then, modulo some data-oscillation
  term we expect to be of higher order, for any $v^\delta \in X^\delta$,
  \[
    \|u - v^\delta\|_X^2 \eqsim \la (\mathbf A_s^{\udelta})^{-1}{\bf r}^\udelta_Y, {\bf r}^\udelta_Y \ra + \|u_0 - \gamma_0 v^\delta\|_H^2,
  \]
  where $\mathbf A_s^{\udelta} := (A_s E_Y^{\udelta} \Xi^{\udelta})(E_Y^{\udelta} \Xi^{\udelta})$ and ${\bf r}^\udelta_Y := (g - B E_X^{\delta} v^\delta)(\Xi^\udelta)$.
  \jw{zouden we in staat zijn om de constantes hidden in $\eqsim$ hierboven, precies te maken? Mss de data oscillation erin?}
\end{prop}
\begin{proof}
  Define $e^\delta := u - E_X^\delta v^\delta$. From~\eqref{eqn:B-lis}, we derive
  \[
    \|e^\delta\|_X^2 \eqsim \|B e^\delta\|_{Y' \times H}^2
    = \|B e^\delta\|_{Y'}^2 + \|\gamma_0 e^\delta\|_H^2 = \|B e^\delta\|_{Y'}^2 + \|u_0 - \gamma_0 v^\delta\|_H^2,
  \]
  and using~\cite[Lemma 3.4]{Stevenson2020a} together with~\eqref{eqn:infinfsup}, we find
  \begin{align*}
    \|B e^\delta\|_{Y'}^2 &= \|e^\delta\|_Y^2 + \|\partial_t e^\delta\|_{Y'}^2 + \|\gamma_T e^\delta\|_H^2 - \|\gamma_0 e^\delta\|_H^2 \\
    &\eqsim \|e^\delta\|_Y^2 + \|\partial_t e^\delta\|_{{Y^\udelta}'}^2 + \|\gamma_T e^\delta\|_H^2 - \|\gamma_0 e^\delta\|_H^2 = \|B e^\delta\|_{{Y^\udelta}'}^2.
  \end{align*}
  The result then follows from
  \begin{align*}
    \|B e^\delta\|_{{Y^\udelta}'}^2 &= \|g - BE_X^\delta v^\delta\|_{{Y^\udelta}'}^2
    = \sup_{\mathclap{0 \not= w \in Y^\udelta}} \frac{[(g - Bv^\delta)(w)]^2}{\|w\|_Y^2} = ~~\sup_{\mathclap{0 \not= {\bf w} \in \R^{\# Y^\udelta}}} ~~\frac{[(g - Bv^\delta)({\bf w}^\top \Xi^\udelta)]^2}{{\bf w}^\top {\bf A}_s^\udelta {\bf w}} \\
    &= \sup_{0 \not= {\bf w} \in \R^{\# Y^\udelta}} \frac{[{\bf w}^\top {\bf r}^\udelta_Y]^2}{{\bf w}^\top {\bf A}_s^\udelta {\bf w}} = \sup_{0 \not= {\bf x} \in \R^{\# Y^\udelta}} \frac{[(({\bf A}_s^\udelta)^{-1/2}{\bf x})^\top {\bf r}^\udelta_Y]^2}{{\bf x}^\top {\bf x}} = \la ({\bf A}_s^\udelta)^{-1} {\bf r}_Y^\udelta, {\bf r}_Y^\udelta \ra. \qedhere
  \end{align*}
\end{proof}
\begin{remark}
  \jw{deze afschatting geldt voro elke $v^\delta \in X^\delta$ en niet alleen voor de oplossing van ons systeem; ik zie twee mogelijk-leuke toepassingen hiervan, namelijk
  \begin{itemize}
    \item het stelt ons in staat om de $X$-norm fout van PCG iterands te plotten
    \item het stelt ons in staat om de $X$-norm fout van (1 component van) de FK19-oplossing te plotten
  \end{itemize}
  is daar iets nuttig aan?}
\end{remark}

@@

Given a $\delta \in \Delta$, and thinking of $f^\delta$ being a quasi-best
approximation to $f$ from $F^\delta$, the difference $f-f^\delta$ is often referred
to as \emph{data-oscillation}.

@@

\subsection{Preconditioning the arising system}
\jw{introduce operator preconditioning}

>>>>

\usepackage{booktabs}
\usepackage{showlabels}
\usepackage[boxed]{algorithm2e}
\usepackage{amsaddr}
\usepackage{babel}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{commath}
\usepackage{colortbl}
\usepackage{dsfont}
\usepackage{enumerate}
\usepackage{esint}
\usepackage[left=0.09\paperwidth, right=0.15\paperwidth, bottom=0.12\paperheight]{geometry}
\usepackage{graphicx}
\usepackage[displaymath, mathlines]{lineno}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{multirow,bigdelim}
\usepackage{url}
\usepackage{pdflscape}
\usepackage{rotating}
%\usepackage[outer]{showlabels}
\usepackage[color=white,linecolor=black, textwidth=0.12\paperwidth]{todonotes}
%\setlength{\parindent}{0pt}
\usepackage{wrapfig}

@@

\usepackage{color}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\providecommand\C{}
\renewcommand{\C}{\mathbb{C}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\K}{\mathcal{K}}
\renewcommand{\H}{\mathcal{H}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\bbD}{\mathbb{D}}
\newcommand{\f}{\varphi}
\newcommand{\e}{\varepsilon}
\newcommand{\eps}{\varepsilon}
\renewcommand{\a}{\alpha}
\renewcommand{\b}{\beta}
\newcommand{\z}{\zeta}
\renewcommand{\O}{\Omega}
\renewcommand{\P}{\mathbb P}
\newcommand{\intRn}{\int_{\R^n}}
\newcommand{\sumi}{\sum_{i=1}^n}
\newcommand{\sumij}{\sum_{i,j=1}^n}
\renewcommand{\d}{\delta}
\newcommand{\p}{\partial}
\DeclareMathOperator{\osc}{osc}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\vol}{vol}
\DeclareMathOperator{\children}{children}
\DeclareMathOperator{\gen}{gen}
\DeclareMathOperator{\cat}{Cat}
\DeclareMathOperator{\Cat}{Cat}
\DeclareMathOperator{\Div}{div}
\DeclareMathOperator{\blockdiag}{blockdiag}
\renewcommand{\div}{\Div}
\DeclareMathOperator{\spann}{span}
\DeclareMathOperator{\diam}{diam}
\DeclareMathOperator{\dist}{dist}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\usepackage{stmaryrd}
\newcommand{\jump}[1]{\left\llbracket #1 \right\rrbracket}
\newcommand{\rhu}{\rightharpoonup}
\newcommand{\T}{\mathcal{T}}
\newcommand{\V}{\mathcal{V}}
\DeclareMathOperator*{\dof}{\#\!DoF}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\Id}{Id}
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\definecolor{ao(english)}{rgb}{0.0, 0.5, 0.0}
\providecommand\note{}
\newcommand{\new}[1]{{\color{ao(english)} #1}}
\renewcommand{\L}{\mathcal L}
\newcommand{\Y}{\mathcal Y}
\newcommand{\Lto}{{L_2(\Omega)}}
\newcommand{\bbT}{\mathbb{T}}
\newcommand{\cL}{\mathcal L}
\newcommand{\Lis}{\cL\mathrm{is}}
\newcommand{\udelta}{{\underline{\delta}}}
\newcommand{\jw}[1]{{\color{red}{JW: #1}}}
\newcommand{\rvv}[1]{{\color{teal}{RvV: #1}}}
\DeclareMathOperator*{\ran}{ran}
\SetKwRepeat{Do}{do}{while}%

@@

\section{Introduction}
Een paar open vragen:
\begin{itemize}
  \item Willen we de algoritmes van \S 5.1, \S 5.2 hierin zetten of citen?
  \item Willen we de details van onze implementatie hier bespreken of willen we die in een (korte?) followup doen?
  \item Wat doen we met secties \S 5.8 5.9, 5.10 uit followup, wat een bewijs geeft van $X$-stability voor een basis die we niet meer gebruiken?

\end{itemize}
\subsection{Notation}
TODO
\section{Problem statement and solution}
Let $V$ and $H$ be separable Hilbert spaces on a spatial domain $\Omega$ such that
$V \hookrightarrow H$ with dense and compact embedding. Identifying $H$ with its
dual, we obtain the Gelfand triple $V \hookrightarrow H \simeq H' \hookrightarrow V'$.

@@

\begin{lemma}
  \label{lem:energy-norm}
  It holds that $\|\cdot\|_X^2=(S\cdot)(\cdot)$.
\end{lemma}
\begin{proof}
  We find
  \[
    \|w\|_X^2=\sup_{0 \neq v_1 \in Y} \frac{(B w)(v_1)}{\|v_1\|_Y^2}+\|\gamma_0 w\|^2
    =\sup_{0 \neq (v_1,v_2) \in Y \times H} \frac{((B w)(v_1)+\la \gamma_0 w,v_2\ra)^2}{\|v_1\|_Y^2+\|v_2\|^2} =(Sw)(w),
  \]
  where the first equality can be found in e.g.~\cite[Thm.~2.1]{Ern2017a}, and,
  when realising that $S$ is also the Schur complement of
  $\begin{bsmallmatrix} A_s & 0 & B\\ 0 & \Id & \gamma_0\\ B' & \gamma_0' & 0\end{bsmallmatrix} \in \cL(Y\times H \times X, Y'\times H'\times X')$, the last one follows from e.g.~\cite[Lemma~2.2]{Kondratyuk2008}.
\end{proof}

@@

Let us first argue why uniform stability for general nonuniform prismatic partitions
is out of reach for our current formulation. \jw{follows from $\dim Y^\delta \lesssim \dim X^\delta$ en
de non-locality van $A_s^{-1}$: gegeven een $X^\delta$ die naar een spacetime-hoekje
gerefined is, moet $Y^\delta$ de gehele time-slab refinen om de inf-sup constante
goed te krijgen. Ik ben het precieze argument
even vergeten, dat staat in een schrift dat op KdV ligt, maar misschien weet Rob dit nog?}

@@

\begin{prop} \label{prop:space-infsup}
  Let $W$ and $\tilde W$ be closed subspaces of $H$ such that there exists a
  (biorthogonal) projector $Q \in \cL(H,H)$ with $\ran Q=W$ and $\ran (\Id -Q)={\tilde W}^\perp$.
  Then, when $W \subset V$, it holds that
  \[
    \gamma:=\inf_{0 \neq \tilde w \in \tilde W}\sup_{0 \neq w \in W}\frac{\la \tilde w,w\ra}{\|\tilde w\|_{V'}\|w\|_V}>0
  \]
  if and only if $Q \in \cL(V,V)$, in which case $\gamma=\|Q\|_{\cL(V,V)}^{-1}$.
\end{prop}
\begin{proof} If $Q\in \cL(V,V)$, then for $\tilde w \in \tilde W$,
$$
\sup_{0 \neq w \in W}\frac{\la \tilde w,w\ra}{\|w\|_V}=\sup_{0 \neq v \in V}\frac{\la \tilde w, v\ra}{\|Qv\|_V}
  \geq \sup_{0 \neq v \in V} \frac{\la \tilde w, v \ra}{\|Q\|_{\cL(V, V)} \|v\|_V} = \|Q\|_{\cL(V,V)}^{-1} \|\tilde w\|_{V'},
$$
or $\gamma \geq \|Q\|_{\cL(V,V)}^{-1}$. If on the other hand $\gamma>0$, then for any $u \in H$,
$$
\gamma \|Q' u\|_{V'} \leq \sup_{0 \neq w \in W}\frac{\la Q'u,w\ra}{\|w\|_V} =
 \sup_{0 \neq w \in W}\frac{\la u,w\ra}{\|w\|_V} \leq \|u\|_{V'},
 $$
so that from $H$ being dense in $V'$, $\|Q\|_{\cL(V,V)}=\|Q'\|_{\cL(V',V')}\leq \gamma^{-1}$.
\end{proof}
\begin{remark}
  In the typical case where $H = L_2(\Omega)$ and $V = H_0^1(\Omega)$ on a bounded
  polytopal domain $\Omega \subset \R^d$, the uniform boundedness of $\|Q\|_{\cL(V,V)}$
  has been demonstrated for families of locally refined partitions, for $d=2$
  including those that are generated by the newest vertex bisection algorithm.
  We refer to \cite{Carstensen2001a,Gaspoz2016}. \jw{Misschien is het ``$H^1$-stability
  of $L^2$-projection" inmiddels wel uitgespeeld. ik vond ook een werk uit
  2014 van de groep van Dirk, en dit werkte voor elke dimensie, maar enkel op
  vreemde fin.elts. Ook een werk uit 2019 van Gaspoz, Heine, Siebert waar het
  2d-NVB geval wordt bekeken voor Lagrange-fin elts tot graad 4.}
\end{remark}

@@

In the spatial axis, we consider the collection $\mathbb T$ of conforming
triangulations generated by NVB from some coarsest triangulation $\T_\perp$,
and define $V_\T = \tilde V_\T$ the space of continuous piecewise linears wrt.~$\T$
that vanish on $\partial \Omega$ such that $X^\delta := \sum_{\lambda \in \vee^\delta_\Sigma} \sigma_\lambda \otimes V_{\T_{\delta, \lambda}}$ satisfies the doubletree constraint.
This means that when $\mu \in \vee^\delta_\Sigma$ is a parent of $\lambda \in \vee^\delta_\Sigma$,
    then $\T_{\delta, \mu}$ is a refinement of $\T_{\delta, \lambda}$, and that $\vee^\delta_\Sigma$ is a subtree of $\vee_\Sigma$, meaning
\begin{itemize}
  \item it contains all wavelets on level 0, and
  \item if it contains a wavelet of level $\lambda \geq 1$, then it also contains
    its parents.
\end{itemize}

@@

We consider spaces $Y^\delta = \tilde Y^\delta$ of the form
$Y^\delta := \sum_{\lambda \in \vee^\delta_\Psi} \psi_\lambda \otimes V_{\T_{\delta, \lambda}}$
for $\T_{\vee^\delta_\Psi, \lambda} \in \mathbb T$, again satisfying the doubletree
constraint. The choice
\begin{equation}
  Y^\delta = \sum_{\lambda \in \vee_\Sigma^\delta} \sum_{\{\mu \in \vee_\Psi\colon |\mu|=|\lambda|,\,
  |\supp \sigma_\lambda \cap \supp \psi_\mu|>0\}} \psi_\mu \otimes V_{\T_{\delta,\lambda}}
  \label{eqn:generate-Ydelta}
\end{equation}
is a doubletree by virtue of $X^\delta$ being one, and for this pair $(X^\delta, Y^\delta)$,
\eqref{eqn:d_tX-in-Y} and \eqref{eqn:inclusion} are valid.

@@

\begin{assum}[$X$-stable splitting of $X^{\udelta} \ominus X^\delta$]
  \jw{is dit een juiste naam voor deze assumption?}
  Assume we have $\Psi^\delta=\{\psi^\delta_\lambda\colon \lambda \in I_\delta^\udelta\} \subset X^\udelta$
  such that $X^\delta + \spann \Psi^\delta=X^\udelta$ and, for some constants $0<m\leq M$,
  for all $z \in X^\delta$ and ${\bf c} := (c_\lambda)_{\lambda \in I_\delta^\udelta} \in \R^{\# I_\delta^\udelta}$, we have
  \begin{equation}
    m^2 \|z+{\bf c}^\top \Psi^\delta\|_X^2
    \leq \|z\|_X^2+\|{\bf c}\|^2
    \leq M^2 \|z+{\bf c}^\top \Psi^\delta\|_X^2.
  \label{eqn:X-stable-splitting}
  \end{equation}
  \label{assum:X-stable-splitting}
\end{assum}
Given $X^\delta = \sum_{\lambda \in \vee_\Sigma^\delta} \sigma_\lambda \otimes V_{\T_{\delta, \lambda}}$, we define
\begin{equation}
  \label{eqn:X-udelta}
  X^\udelta := \sum_{\lambda \in \vee_\Sigma^\delta} \left[\sum_{\mu \in \children(\lambda)} \sigma_\mu \otimes V_{\T_{\delta, \lambda}} +  \sigma_\lambda \otimes V_{\T^{++}_{\delta, \lambda}} \right]
\end{equation}
where $\T^{++}$ is found from $\T$ by two uniform bisections. This definition of $X^\udelta$ defines some nodes twice.\jw{slechte zin maar die "+" maar het dus niet-uniek; sommige nodes komen van "een pijltje omhoog" en "een pijltje naar rechts"}
\begin{remark}
  When $X^\delta$ is a sparse grid until level $2L$ \jw{de definitie hiervan wordt pas in \S 6 gegeven}, then $X^\udelta$ is a sparse grid until level $2(L+1)$.
\end{remark}
\begin{remark}
  A suitable basis is not currently known to us. However, in Appendix~\ref{sec:PsiDelta}, we will construct
  such $\Psi^\delta$ with constants $m$ and $M$ that deteriorate with increasing $|\lambda|$.
  \jw{is die constructie iets dat we misschien in een appendix willen zetten?}
\end{remark}

@@

This error estimator will drive our adaptive loop. Even without Assumption~\ref{assum:X-stable-splitting} however, we derive the following efficient and reliable \emph{global estimator}.
\begin{prop}
  \label{prop:global-apost}
  Let $\Xi^{\udelta}$ be a basis for $Y^{\udelta}$. Then, modulo some data-oscillation
  term we expect to be of higher order, for any $v^\delta \in X^\delta$,
  \[
    \|u - v^\delta\|_X^2 \eqsim \la (\mathbf A_s^{\udelta})^{-1}{\bf r}^\udelta_Y, {\bf r}^\udelta_Y \ra + \|u_0 - \gamma_0 v^\delta\|^2,
  \]
  where $\mathbf A_s^{\udelta} := (A_s E_Y^{\udelta} \Xi^{\udelta})(E_Y^{\udelta} \Xi^{\udelta})$ and ${\bf r}^\udelta_Y := (g - B E_X^{\delta} v^\delta)(\Xi^\udelta)$.
  \jw{zouden we in staat zijn om de constantes hidden in $\eqsim$ hierboven, precies te maken? Mss de data oscillation erin?}
\end{prop}
\begin{proof}
  Define $e^\delta := u - E_X^\delta v^\delta$. From~\eqref{eqn:B-lis}, we derive
  \[
    \|e^\delta\|_X^2 \eqsim \|B e^\delta\|_{Y' \times H}^2
    = \|B e^\delta\|_{Y'}^2 + \|\gamma_0 e^\delta\|^2 = \|B e^\delta\|_{Y'}^2 + \|u_0 - \gamma_0 v^\delta\|^2,
  \]
  and using~\cite[Lemma 3.4]{Stevenson2020a} together with~\eqref{eqn:infinfsup}, we find
  \begin{align*}
    \|B e^\delta\|_{Y'}^2 &= \|e^\delta\|_Y^2 + \|\partial_t e^\delta\|_{Y'}^2 + \|\gamma_T e^\delta\|^2 - \|\gamma_0 e^\delta\|^2 \\
    &\eqsim \|e^\delta\|_Y^2 + \|\partial_t e^\delta\|_{{Y^\udelta}'}^2 + \|\gamma_T e^\delta\|^2 - \|\gamma_0 e^\delta\|^2 = \|B e^\delta\|_{{Y^\udelta}'}^2.
  \end{align*}
  The result then follows from
  \begin{align*}
    \|B e^\delta\|_{{Y^\udelta}'}^2 &= \|g - BE_X^\delta v^\delta\|_{{Y^\udelta}'}^2
    = \sup_{\mathclap{0 \not= w \in Y^\udelta}} \frac{[(g - Bv^\delta)(w)]^2}{\|w\|_Y^2} = ~~\sup_{\mathclap{0 \not= {\bf w} \in \R^{\dim Y^\udelta}}} ~~\frac{[(g - Bv^\delta)({\bf w}^\top \Xi^\udelta)]^2}{{\bf w}^\top {\bf A}_s^\udelta {\bf w}} \\
    &= \sup_{0 \not= {\bf w} \in \R^{\dim Y^\udelta}} \frac{[{\bf w}^\top {\bf r}^\udelta_Y]^2}{{\bf w}^\top {\bf A}_s^\udelta {\bf w}} = \sup_{0 \not= {\bf x} \in \R^{\dim Y^\udelta}} \frac{[(({\bf A}_s^\udelta)^{-1/2}{\bf x})^\top {\bf r}^\udelta_Y]^2}{{\bf x}^\top {\bf x}} = \la ({\bf A}_s^\udelta)^{-1} {\bf r}_Y^\udelta, {\bf r}_Y^\udelta \ra. \qedhere
  \end{align*}
\end{proof}
\begin{remark}
  \jw{deze afschatting geldt voro elke $v^\delta \in X^\delta$ en niet alleen voor de oplossing van ons systeem; ik zie twee mogelijk-leuke toepassingen hiervan, namelijk
  \begin{itemize}
    \item het stelt ons in staat om de $X$-norm fout van PCG iterands te plotten
    \item het stelt ons in staat om de $X$-norm fout van (1 component van) de FK19-oplossing te plotten
  \end{itemize}
  is daar iets nuttig aan?}
\end{remark}

@@

Given a $\delta \in \Delta$, and thinking of $f^\delta$ being a quasi-best
approximation to $f$ from $F^\delta$, the difference $f-f^\delta$ is often referred
to as \emph{data-oscillation}.
\jw{hier iets over de specificatie van $F^\delta$ en $f^\delta$ zoals \S 5.8 uit followup}

@@

\subsection{Preconditioning the arising system}
\jw{introduce operator preconditioning}. \jw{onze norm-inducing bilinear forms lijken wat op die van [arXiv:1212.6037]. Maar onze $D_X$-operator is mooier, door dat inzicht van [PW12]}
